Succinct representations for abstract interpretation

### 2011-11-02 papers/PLDI2012/implicitization.tex
suppression commentaires + correction bibtex

\MM{Il faudra peut-être une section ``détails d'implem'' où mettre
  l'histoire des BDDs pour P. Je laisse ici faute de mieux (peut-être
  que la section~\ref{sec:analysis-algorithm} ferait l'affaire ?),
  mais ça distrait un peu le lecteur}

@@

\MM{Pas sûr que ça soit nécessaire, mais vu qu'on a eu une version
  fausse avant, c'est peut-être bien de dire comment on est arrivés à
  la meilleure version.}
The order of steps is important: narrowing has to be performed before
adding new paths, or some spurious new paths would be added to $P$,
and starting with the addition of new paths avoids wasting time
doing the ascending iterations on an empty graph.

@@

The $\sigma_i$ functions take as parameters the index of the starting abstract
value, and the path we focus on. If we dynamically construct these functions
during the analysis, we should store the results of the $\sigma_i$ functions into a
compact representation, such as Algebraic Decision Diagrams.\MM{Du
  coup, j'aurais intégré cette phrase au paragraphe suivant. C'est des
  ADD que tu utilises finalement ?}
Note that $\sigma_i(j,k)$ will be computed only if needed.

We experimented with a heuristic
of dynamic construction of the $\sigma_i$ functions, adapted from
\citet{DBLP:conf/pldi/GulwaniZ10}. 
\DM{Cette heuristique, je croyais que c'était nous qui l'avions trouvée,
et qu'elle n'était pas dans Gulwani?}
\JH{Non. Gulwani fait la même chose en un peu plus complique (voir commentaire
suivant)}
\MM{Du coup, la contribution~\ref{contr:disjunctive} devrait être
  affaiblie un peu je suppose. Gulwani fait aussi du SMT-solving ?}
\DM{Non, Gulwani énumère tous les chemins un par un à tous les coups.}
We start with one single
disjunct per control point ($m_i=1$ for each $i, p_i \in P_R$) 
and empty maps $\sigma_i$. We also define $\delta_i = 1$ for each $i, p_i \in
P_R$.
$M$ denotes an upper bound on the number of disjuncts per control point.
$\sigma_i(j,k)$ is then constructed on the fly only when needed.
When $\sigma_i(j,k)$ needs to be computed, we first compute the image of the
abstract value $X_{i,j}$ by the path indexed by $k$, and try to find an existing
disjunct of index $j'$ so that the least upper bound of the two abstract values is exactly their union (this can be tested using SMT-solving).
If such an index exists, then we set $\sigma_i(j,k) = j'$.
Otherwise:
\begin{itemize}[itemsep=-1\parsep,topsep=0pt]
	\item if $m_i < M$, we increase $m_i$ by $1$ and define $\sigma_i(j,k) = m_i$
	\item if $m_i = M$, we define $\sigma_i(j,k) = M$ 
		\JH{A la place, Gulwani cherche l'invariant qui a le plus de contraintes
		d'egalite en commun pour definir son $sigma$. Nous, on le met dans M, pour simplifier}
\end{itemize}

@@

\MM{On peut peut-être citer le nom de l'outil (on garde PAGAI ?) ici.
  L'intérêt d'avoir un nom, c'est que les gens peuvent en parler plus
  facilement, et c'est plus facile de retrouver le papier origine d'un
  outil quand il cite son nom}
\DM{Dans la version finale, oui, mais pas dans une soumission anonymisée... suffit de Googlifier pagai sur imag.fr}

Our tool operates over LLVM bitcode \citep{LLVM_langref,Lattner:2004:LCF:977395.977673}, which is a target for several compilers, most notably Clang (supporting C and C++) and llvm-gcc (supporting C, C++, Fortran and Ada).
Abstract domains are provided by the APRON library \citep{DBLP:conf/cav/JeannetM09}, and include convex polyhedra (from the builtin Polka ``PK'' library), octagons, intervals, and linear congruences.
For SMT-solving, our analyzer uses Yices
\cite{DBLP:conf/cav/DutertreM06} or Microsoft Z3\cite{DBLP:conf/tacas/MouraB08}.

>>>>

@@

The order of steps is important: narrowing has to be performed before
adding new paths, or some spurious new paths would be added to $P$,
and starting with the addition of new paths avoids wasting time
doing the ascending iterations on an empty graph.

@@

We experimented with a heuristic
of dynamic construction of the $\sigma_i$ functions, adapted from
\citet{DBLP:conf/pldi/GulwaniZ10}. 
%\DM{Cette heuristique, je croyais que c'était nous qui l'avions trouvée,
%et qu'elle n'était pas dans Gulwani?}
%\JH{Non. Gulwani fait la même chose en un peu plus complique (voir commentaire
%suivant)}
%\MM{Du coup, la contribution~\ref{contr:disjunctive} devrait être
%  affaiblie un peu je suppose. Gulwani fait aussi du SMT-solving ?}
%\DM{Non, Gulwani énumère tous les chemins un par un à tous les coups.}

For each control point $p_i \in P_R$, we start with one single disjunct
($m_i = 1$) and define $\delta_i = 1$.
$M$ denotes an upper bound on the number of disjuncts per control point.

The $\sigma_i$ functions take as parameters the index of the starting abstract
value, and the path we focus on. 
Since we dynamically construct these functions
during the analysis, we store their already comupted image into a
compact representation, such as Algebraic Decision Diagrams.
$\sigma_i(j,k)$ is then constructed on the fly only when needed, and computed
only once.
When the value of $\sigma_i(j,k)$ is required but undefined, we first compute
the image of the
abstract value $X_{i,j}$ by the path indexed by $k$, and try to find an existing
disjunct of index $j'$ so that the least upper bound of the two abstract values is exactly their union (this can be tested using SMT-solving).
If such an index exists, then we set $\sigma_i(j,k) = j'$.
Otherwise:
\begin{itemize}[itemsep=-1\parsep,topsep=0pt]
	\item if $m_i < M$, we increase $m_i$ by $1$ and define $\sigma_i(j,k) = m_i$
	\item if $m_i = M$, we define $\sigma_i(j,k) = M$ 
		%\JH{A la place, Gulwani cherche l'invariant qui a le plus de contraintes
		%d'egalite en commun pour definir son $sigma$. Nous, on le met dans M, pour simplifier}
\end{itemize}

@@

%\MM{On peut peut-être citer le nom de l'outil (on garde PAGAI ?) ici.
%  L'intérêt d'avoir un nom, c'est que les gens peuvent en parler plus
%  facilement, et c'est plus facile de retrouver le papier origine d'un
%  outil quand il cite son nom}
%\DM{Dans la version finale, oui, mais pas dans une soumission anonymisée... suffit de Googlifier pagai sur imag.fr}
%
Our tool operates over LLVM bitcode \citep{LLVM_langref,Lattner:2004:LCF:977395.977673}, which is a target for several compilers, most notably Clang (supporting C and C++) and llvm-gcc (supporting C, C++, Fortran and Ada).
Abstract domains are provided by the APRON library \citep{DBLP:conf/cav/JeannetM09}, and include convex polyhedra (from the builtin Polka ``PK'' library), octagons, intervals, and linear congruences.
For SMT-solving, our analyzer uses Yices
\cite{DBLP:conf/cav/DutertreM06} or Microsoft Z3\cite{DBLP:conf/tacas/MouraB08}.
